Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision (\YellowPaperVersionNumber{})}}

### 2017-10-19 Paper.tex

Ethereum, taken as a whole, can be viewed as a transaction-based state machine: we begin with a genesis state and incrementally execute transactions to morph it into some final state. It is this final state which we accept as the canonical ``version'' of the world of Ethereum. The state can include such information as account balances, reputations, trust arrangements, data pertaining to information of the physical world; in short, anything that can currently be represented by a computer is admissible. Transactions thus represent a valid arc between two states; the `valid' part is important---there exist far more invalid state changes than valid state changes. Invalid state changes might, \eg, be things such as reducing an account balance without an equal and opposite increase elsewhere. A valid state transition is one which comes about through a transaction. Formally:\hypertarget{Upsilon}{}
\begin{equation}
\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_\mathrm{t}, T)
\end{equation}

@@

Formally, we expand to:
\begin{eqnarray}
\boldsymbol{\sigma}_{t+1} & \equiv & \hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}_\mathrm{t}, B) \\
B & \equiv & (..., (T_0, T_1, ...) ) \\
\Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Omega}{\Omega}(B, \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
\end{eqnarray}

@@

Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction. This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg  \hyperlink{transaction nonce}{$T_\mathrm{n}$, denotes the nonce of said transaction}. The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.

@@

Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\boldsymbol{\mu}_\mathrm{s}[0]$ denotes the first item on the machine's stack. For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\boldsymbol{\mu}_\mathrm{m}[0..31]$ denotes the first 32 items of the machine's memory.

@@

\begin{description}
\hypertarget{account nonce}{}
\item[nonce] A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account. For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_\mathrm{n}$.
\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_\mathrm{b}$.
\item[storageRoot] A 256-bit hash of the root node of a Merkle Patricia tree that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from the Keccak 256-bit hash of the  256-bit integer keys to the RLP-encoded 256-bit integer values. The hash is formally denoted $\boldsymbol{\sigma}[a]_\mathrm{s}$.
\item[codeHash] The hash of the EVM code of this account---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction. All such code fragments are contained in the state database under their corresponding hashes for later retrieval. This hash is formally denoted $\boldsymbol{\sigma}[a]_\mathrm{c}$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_\mathrm{c}$.
\end{description}

Since I typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, I define a convenient equivalence:
\begin{equation}
\texttt{\small TRIE}\big(L_I^*(\boldsymbol{\sigma}[a]_\mathrm{s})\big) \equiv \boldsymbol{\sigma}[a]_\mathrm{s}
\end{equation}

@@

It shall be understood that $\boldsymbol{\sigma}[a]_\mathrm{s}$ is not a `physical' member of the account and does not contribute to its later serialisation.

If the \textbf{codeHash} field is the Keccak-256 hash of the empty string, i.e. $\boldsymbol{\sigma}[a]_\mathrm{c} = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.

Thus we may define a world-state collapse function $L_S$:
\begin{equation}
L_S(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
\end{equation}
where
\begin{equation}
p(a) \equiv  \big(\texttt{\small KEC}(a), \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_\mathrm{n}, \boldsymbol{\sigma}[a]_\mathrm{b}, \boldsymbol{\sigma}[a]_\mathrm{s}, \boldsymbol{\sigma}[a]_\mathrm{c}) \big) \big)
\end{equation}

This function, $L_S$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
\begin{equation}
\forall a: \boldsymbol{\sigma}[a] = \varnothing \; \vee \; (a \in \mathbb{B}_{20} \; \wedge \; v(\boldsymbol{\sigma}[a]))
\end{equation}
\hypertarget{v x}{where $v$ is the account validity function:
\begin{equation}
\quad v(x) \equiv x_\mathrm{n} \in \mathbb{P}_{256} \wedge x_\mathrm{b} \in \mathbb{P}_{256} \wedge x_\mathrm{s} \in \mathbb{B}_{32} \wedge x_\mathrm{c} \in \mathbb{B}_{32}
}
\end{equation}

@@

\begin{description}
\hypertarget{T n}{
\item[nonce] \label{itm:nonce} A scalar value equal to the number of transactions sent by the sender; formally $T_\mathrm{n}$.}\hypertarget{T p}{
\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_\mathrm{p}$.}\hypertarget{T g}{
\item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction. This is paid up-front, before any computation is done and may not be increased later; formally $T_\mathrm{g}$.}\hypertarget{T t}{
\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_\mathrm{t}$.}\hypertarget{T v}{
\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_\mathrm{v}$.}\hypertarget{T w}{
\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally {$T_\mathrm{w}$}}, \hyperlink{T r}{$T_\mathrm{r}$} and \hyperlink{T s}{$T_\mathrm{s}$}, respectively. This is expanded in Appendix \ref{app:signing}.
\end{description}

@@

\begin{description}
\item[init] An unlimited size byte array specifying the EVM-code for the account initialisation procedure, formally $T_\mathrm{i}$.
\end{description}

@@

\begin{description}
\item[data] An unlimited size byte array specifying the input data of the message call, formally $T_\mathrm{d}$.
\end{description}

\begin{equation}
L_\mathrm{T}(T) \equiv \begin{cases}
(T_\mathrm{n}, T_\mathrm{p}, T_\mathrm{g}, T_\mathrm{t}, \hyperlink{T v}{T_\mathrm{v}}, T_\mathrm{i}, \hyperlink{T w}{T_\mathrm{w}}, T_\mathrm{r}, T_\mathrm{s}) & \text{if} \; T_\mathrm{t} = \varnothing\\
(T_\mathrm{n}, T_\mathrm{p}, T_\mathrm{g}, T_\mathrm{t}, \hyperlink{T v}{T_\mathrm{v}}, T_\mathrm{d}, \hyperlink{T w}{T_\mathrm{w}}, T_\mathrm{r}, T_\mathrm{s}) & \text{otherwise}
\end{cases}
\end{equation}

Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_\mathrm{i}$ and $T_\mathrm{d}$.
\begin{equation}
\begin{array}[t]{lclclc}
T_\mathrm{n} \in \mathbb{P}_{256} & \wedge & \hyperlink{T v}{T_\mathrm{v}} \in \mathbb{P}_{256} & \wedge & T_\mathrm{p} \in \mathbb{P}_{256} & \wedge \\
T_\mathrm{g} \in \mathbb{P}_{256} & \wedge & \hyperlink{T w}{T_\mathrm{w}} \in \mathbb{P}_5 & \wedge & T_\mathrm{r} \in \mathbb{P}_{256} & \wedge \\
T_\mathrm{s} \in \mathbb{P}_{256} & \wedge & T_\mathrm{d} \in \mathbb{B} & \wedge & T_\mathrm{i} \in \mathbb{B}
\end{array}
\end{equation}
where
\begin{equation}
\mathbb{P}_\mathrm{n} = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
\end{equation}

The address hash $T_\mathrm{t}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte sequence and thus the member of $\mathbb{B}_0$:
\begin{equation}
T_\mathrm{t} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_\mathrm{t} \neq \varnothing \\
\mathbb{B}_{0} & \text{otherwise}\end{cases}
\end{equation}

@@

\begin{description}\hypertarget{H p}{
\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_\mathrm{p}$.}
\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $\hyperlink{H o}{H_\mathrm{o}}$.\hypertarget{H c}{
\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block shall be transferred; formally $H_\mathrm{c}$.}
\item[stateRoot] The Keccak 256-bit hash of the root node of the state \hyperlink{trie}{trie}, after all transactions are executed and finalisations applied; formally \hyperlink{H r}{$H_\mathrm{r}$}.
\item[transactionsRoot] The Keccak 256-bit hash of the root node of the \hyperlink{trie}{trie} structure populated with each transaction in the transactions list portion of the block; formally \hyperlink{H t}{$H_\mathrm{t}$}.
\item[receiptsRoot] The Keccak 256-bit hash of the root node of the \hyperlink{trie}{trie} structure populated with the receipts of each transaction in the transactions list portion of the block; formally \hyperlink{H e}{$H_\mathrm{e}$}.
\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally \hyperlink{H b}{$H_\mathrm{b}.$}\hyperlink{H d}{
\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_\mathrm{d}$}.\hypertarget{H i}{
\item[number] A scalar value equal to the number of ancestor blocks. The \hyperlink{GenesisBlock}{genesis block} has a number of zero; formally $H_\mathrm{i}$.}\hyperlink{H l}{
\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_\mathrm{l}$.}\hypertarget{H g}{
\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_\mathrm{g}$.}\hypertarget{H s}{
\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally {$H_\mathrm{s}$}.}\hypertarget{H x}{
\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_\mathrm{x}$.}\hypertarget{H m}{
\item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_\mathrm{m}$.}
\hypertarget{block nonce}{\hypertarget{H n}{
\item[nonce] A 64-bit hash which proves combined with the mix-hash that a sufficient amount of computation has been carried out on this block; formally $H_\mathrm{n}$.}}
\end{description}
\hypertarget{B U}{
The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_\mathrm{U}$, and a series of the transactions, $B_\mathrm{T}$. Formally, we can refer to a block $B$:
\begin{equation}
B \equiv (B_H, B_\mathrm{T}, B_\mathrm{U})
\end{equation}}

@@

In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
Each receipt, denoted $B_\mathrm{R}[i]$ for the $i$th transaction, is placed in an index-keyed \hyperlink{trie}{trie} and the root recorded in the header as \hyperlink{H e}{$H_\mathrm{e}$}.

\hypertarget{R}{}
The transaction receipt, $R$, is a tuple of four items \hypertarget{R sigma}{comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$}, \hypertarget{R u}{the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_\mathrm{u}$}, the set of logs created through execution of the transaction, \hyperlink{R l}{$R_\mathrm{l}$} and the Bloom filter composed from information in those logs, \hyperlink{R b}{$R_\mathrm{b}$}:
\begin{equation}
R \equiv (R_{\boldsymbol{\sigma}}, R_\mathrm{u}, \hyperlink{R b}{R_\mathrm{b}}, \hyperlink{R l}{R_\mathrm{l}})
\end{equation}

The function $L_R$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
\begin{equation}
L_R(R) \equiv (\mathtt{\small \hyperlink{trie}{TRIE}}(L_S(R_{\boldsymbol{\sigma}})), \hyperlink{R u}{R_\mathrm{u}}, \hyperlink{R b}{R_\mathrm{b}}, \hyperlink{R l}{R_\mathrm{l}})
\end{equation}
thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a \hyperlink{trie}{trie} structure, the root of which forms the first item.

We assert $\hyperlink{R u}{R_\mathrm{u}}$, the cumulative gas used is a positive integer and that the logs Bloom, \hyperlink{R b}{$R_\mathrm{b}$}, is a hash of size 2048 bits (256 bytes):
\begin{equation}
\hyperlink{R u}{R_\mathrm{u}} \in \mathbb{P} \quad \wedge \quad \hyperlink{R b}{R_\mathrm{b}} \in \mathbb{B}_{256}
\end{equation}

%Notably $B_\mathrm{T}$ does not get serialised into the block by the block preparation function $L_B$; it is merely a convenience equivalence.

The log entries, $\hyperlink{R l}{R_\mathrm{l}}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$. A log entry, $O$, is a tuple of a logger's address, $O_\mathrm{a}$, a series of 32-bytes log topics, $O_\mathrm{t}$ and some number of bytes of data, $O_\mathrm{d}$:
\begin{equation}
O \equiv (O_\mathrm{a}, ({O_\mathrm{t}}_0, {O_\mathrm{t}}_1, ...), O_\mathrm{d})
\end{equation}
\begin{equation}
O_\mathrm{a} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathrm{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathrm{d} \in \mathbb{B}
\end{equation}

We define the Bloom filter function, $M$, to reduce a log entry into a single 256-byte hash:
\begin{equation}
M(O) \equiv \hyperlink{bigvee}{\bigvee}_{t \in \{O_\mathrm{a}\} \cup O_\mathrm{t}} \big( M_{3:2048}(t) \big)
\end{equation}

@@

Furthermore:
\begin{equation}
\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_H)_H}_\mathrm{r}
\end{equation}

@@

The values stemming from the computation of transactions, specifically the \hyperlink{Transaction Receipt}{transaction receipts, $B_\mathrm{R}$}, and that defined through the transaction's \hyperlink{Pi}{state-accumulation function, $\Pi$}, are formalised later in section \ref{sec:statenoncevalidation}.

@@

We define $P(B_H)$ to be the parent block of $B$, formally:
\begin{equation}
P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_H)) = \hyperlink{H p}{H_\mathrm{p}}
\end{equation}

The block number is the parent's block number incremented by one:
\begin{equation}
\hyperlink{H i}{H_\mathrm{i}} \equiv {{P(H)_H}_\mathrm{i}} + 1
\end{equation}

@@

The nonce of a block, $\hyperlink{H n}{H_\mathrm{n}}$, must satisfy the relations:
\begin{equation}
n \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathrm{d}}} \quad \wedge \quad m = \hyperlink{H m}{H_\mathrm{m}}
\end{equation}
with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, \hyperlink{H n}{H_\mathrm{n}}, \mathbf{d})$.

>>>>

Ethereum, taken as a whole, can be viewed as a transaction-based state machine: we begin with a genesis state and incrementally execute transactions to morph it into some final state. It is this final state which we accept as the canonical ``version'' of the world of Ethereum. The state can include such information as account balances, reputations, trust arrangements, data pertaining to information of the physical world; in short, anything that can currently be represented by a computer is admissible. Transactions thus represent a valid arc between two states; the `valid' part is important---there exist far more invalid state changes than valid state changes. Invalid state changes might, \eg, be things such as reducing an account balance without an equal and opposite increase elsewhere. A valid state transition is one which comes about through a transaction. Formally:\hypertarget{Upsilon}{}
\begin{equation}
\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_{\mathrm{t}}, T)
\end{equation}

@@

Formally, we expand to:
\begin{eqnarray}
\boldsymbol{\sigma}_{t+1} & \equiv & \hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}_{\mathrm{t}}, B) \\
B & \equiv & (..., (T_0, T_1, ...) ) \\
\Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Omega}{\Omega}(B, \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
\end{eqnarray}

@@

Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction. This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg  \hyperlink{transaction nonce}{$T_{\mathrm{n}}$, denotes the nonce of said transaction}. The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.

@@

Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\boldsymbol{\mu}_{\mathrm{s}}[0]$ denotes the first item on the machine's stack. For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\boldsymbol{\mu}_{\mathrm{m}}[0..31]$ denotes the first 32 items of the machine's memory.

@@

\begin{description}
\hypertarget{account nonce}{}
\item[nonce] A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account. For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{n}}$.
\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{b}}$.
\item[storageRoot] A 256-bit hash of the root node of a Merkle Patricia tree that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from the Keccak 256-bit hash of the  256-bit integer keys to the RLP-encoded 256-bit integer values. The hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{s}}$.
\item[codeHash] The hash of the EVM code of this account---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction. All such code fragments are contained in the state database under their corresponding hashes for later retrieval. This hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{c}}$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_{\mathrm{c}}$.
\end{description}

Since I typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, I define a convenient equivalence:
\begin{equation}
\texttt{\small TRIE}\big(L_I^*(\boldsymbol{\sigma}[a]_{\mathrm{s}})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
\end{equation}

@@

It shall be understood that $\boldsymbol{\sigma}[a]_{\mathrm{s}}$ is not a `physical' member of the account and does not contribute to its later serialisation.

If the \textbf{codeHash} field is the Keccak-256 hash of the empty string, i.e. $\boldsymbol{\sigma}[a]_{\mathrm{c}} = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.

Thus we may define a world-state collapse function $L_S$:
\begin{equation}
L_S(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
\end{equation}
where
\begin{equation}
p(a) \equiv  \big(\texttt{\small KEC}(a), \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_{\mathrm{n}}, \boldsymbol{\sigma}[a]_{\mathrm{b}}, \boldsymbol{\sigma}[a]_{\mathrm{s}}, \boldsymbol{\sigma}[a]_{\mathrm{c}}) \big) \big)
\end{equation}

This function, $L_S$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
\begin{equation}
\forall a: \boldsymbol{\sigma}[a] = \varnothing \; \vee \; (a \in \mathbb{B}_{20} \; \wedge \; v(\boldsymbol{\sigma}[a]))
\end{equation}
\hypertarget{v x}{where $v$ is the account validity function:
\begin{equation}
\quad v(x) \equiv x_{\mathrm{n}} \in \mathbb{P}_{256} \wedge x_{\mathrm{b}} \in \mathbb{P}_{256} \wedge x_{\mathrm{s}} \in \mathbb{B}_{32} \wedge x_{\mathrm{c}} \in \mathbb{B}_{32}
}
\end{equation}

@@

\begin{description}
\hypertarget{T n}{
\item[nonce] \label{itm:nonce} A scalar value equal to the number of transactions sent by the sender; formally $T_{\mathrm{n}}$.}\hypertarget{T p}{
\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{p}}$.}\hypertarget{T g}{
\item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction. This is paid up-front, before any computation is done and may not be increased later; formally $T_{\mathrm{g}}$.}\hypertarget{T t}{
\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_{\mathrm{t}}$.}\hypertarget{T v}{
\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_{\mathrm{v}}$.}\hypertarget{T w}{
\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally {$T_{\mathrm{w}}$}}, \hyperlink{T r}{$T_{\mathrm{r}}$} and \hyperlink{T s}{$T_{\mathrm{s}}$}, respectively. This is expanded in Appendix \ref{app:signing}.
\end{description}

@@

\begin{description}
\item[init] An unlimited size byte array specifying the EVM-code for the account initialisation procedure, formally $T_{\mathrm{i}}$.
\end{description}

@@

\begin{description}
\item[data] An unlimited size byte array specifying the input data of the message call, formally $T_{\mathrm{d}}$.
\end{description}

\begin{equation}
L_{\mathrm{t}}(T) \equiv \begin{cases}
(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, \hyperlink{T v}{T_{\mathrm{v}}}, T_{\mathrm{i}}, \hyperlink{T w}{T_{\mathrm{w}}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{t}} = \varnothing\\
(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, \hyperlink{T v}{T_{\mathrm{v}}}, T_{\mathrm{d}}, \hyperlink{T w}{T_{\mathrm{w}}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{otherwise}
\end{cases}
\end{equation}

Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_{\mathrm{i}}$ and $T_{\mathrm{d}}$.
\begin{equation}
\begin{array}[t]{lclclc}
T_{\mathrm{n}} \in \mathbb{P}_{256} & \wedge & \hyperlink{T v}{T_{\mathrm{v}}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{p}} \in \mathbb{P}_{256} & \wedge \\
T_{\mathrm{g}} \in \mathbb{P}_{256} & \wedge & \hyperlink{T w}{T_{\mathrm{w}}} \in \mathbb{P}_5 & \wedge & T_{\mathrm{r}} \in \mathbb{P}_{256} & \wedge \\
T_{\mathrm{s}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{d}} \in \mathbb{B} & \wedge & T_{\mathrm{i}} \in \mathbb{B}
\end{array}
\end{equation}
where
\begin{equation}
\mathbb{P}_{\mathrm{n}} = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
\end{equation}

The address hash $T_{\mathrm{t}}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte sequence and thus the member of $\mathbb{B}_0$:
\begin{equation}
T_{\mathrm{t}} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_{\mathrm{t}} \neq \varnothing \\
\mathbb{B}_{0} & \text{otherwise}\end{cases}
\end{equation}

@@

\begin{description}\hypertarget{H p}{
\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.}
\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $\hyperlink{H o}{H_{\mathrm{o}}}$.\hypertarget{H c}{
\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block shall be transferred; formally $H_{\mathrm{c}}$.}
\item[stateRoot] The Keccak 256-bit hash of the root node of the state \hyperlink{trie}{trie}, after all transactions are executed and finalisations applied; formally \hyperlink{H r}{$H_{\mathrm{r}}$}.
\item[transactionsRoot] The Keccak 256-bit hash of the root node of the \hyperlink{trie}{trie} structure populated with each transaction in the transactions list portion of the block; formally \hyperlink{H t}{$H_{\mathrm{t}}$}.
\item[receiptsRoot] The Keccak 256-bit hash of the root node of the \hyperlink{trie}{trie} structure populated with the receipts of each transaction in the transactions list portion of the block; formally \hyperlink{H e}{$H_{\mathrm{e}}$}.
\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally \hyperlink{H b}{$H_{\mathrm{b}}.$}\hyperlink{H d}{
\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_{\mathrm{d}}$}.\hypertarget{H i}{
\item[number] A scalar value equal to the number of ancestor blocks. The \hyperlink{GenesisBlock}{genesis block} has a number of zero; formally $H_{\mathrm{i}}$.}\hyperlink{H l}{
\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.}\hypertarget{H g}{
\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.}\hypertarget{H s}{
\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally {$H_{\mathrm{s}}$}.}\hypertarget{H x}{
\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.}\hypertarget{H m}{
\item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.}
\hypertarget{block nonce}{\hypertarget{H n}{
\item[nonce] A 64-bit hash which proves combined with the mix-hash that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.}}
\end{description}
\hypertarget{B U}{
The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_{\mathrm{u}}$, and a series of the transactions, $B_{\mathrm{t}}$. Formally, we can refer to a block $B$:
\begin{equation}
B \equiv (B_H, B_{\mathrm{t}}, B_{\mathrm{u}})
\end{equation}}

@@

In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
Each receipt, denoted $B_{\mathrm{r}}[i]$ for the $i$th transaction, is placed in an index-keyed \hyperlink{trie}{trie} and the root recorded in the header as \hyperlink{H e}{$H_{\mathrm{e}}$}.

\hypertarget{R}{}
The transaction receipt, $R$, is a tuple of four items \hypertarget{R sigma}{comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$}, \hypertarget{R u}{the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$}, the set of logs created through execution of the transaction, \hyperlink{R l}{$R_{\mathrm{l}}$} and the Bloom filter composed from information in those logs, \hyperlink{R b}{$R_{\mathrm{b}}$}:
\begin{equation}
R \equiv (R_{\boldsymbol{\sigma}}, R_{\mathrm{u}}, \hyperlink{R b}{R_{\mathrm{b}}}, \hyperlink{R l}{R_{\mathrm{l}}})
\end{equation}

The function $L_R$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
\begin{equation}
L_R(R) \equiv (\mathtt{\small \hyperlink{trie}{TRIE}}(L_S(R_{\boldsymbol{\sigma}})), \hyperlink{R u}{R_{\mathrm{u}}}, \hyperlink{R b}{R_{\mathrm{b}}}, \hyperlink{R l}{R_{\mathrm{l}}})
\end{equation}
thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a \hyperlink{trie}{trie} structure, the root of which forms the first item.

We assert $\hyperlink{R u}{R_{\mathrm{u}}}$, the cumulative gas used is a positive integer and that the logs Bloom, \hyperlink{R b}{$R_{\mathrm{b}}$}, is a hash of size 2048 bits (256 bytes):
\begin{equation}
\hyperlink{R u}{R_{\mathrm{u}}} \in \mathbb{P} \quad \wedge \quad \hyperlink{R b}{R_{\mathrm{b}}} \in \mathbb{B}_{256}
\end{equation}

%Notably $B_{\mathrm{t}}$ does not get serialised into the block by the block preparation function $L_B$; it is merely a convenience equivalence.

The log entries, $\hyperlink{R l}{R_{\mathrm{l}}}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$. A log entry, $O$, is a tuple of a logger's address, $O_{\mathrm{a}}$, a series of 32-bytes log topics, $O_{\mathrm{t}}$ and some number of bytes of data, $O_{\mathrm{d}}$:
\begin{equation}
O \equiv (O_{\mathrm{a}}, ({O_{\mathrm{t}}}_0, {O_{\mathrm{t}}}_1, ...), O_{\mathrm{d}})
\end{equation}
\begin{equation}
O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_{\mathrm{t}}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_{\mathrm{d}} \in \mathbb{B}
\end{equation}

We define the Bloom filter function, $M$, to reduce a log entry into a single 256-byte hash:
\begin{equation}
M(O) \equiv \hyperlink{bigvee}{\bigvee}_{t \in \{O_{\mathrm{a}}\} \cup O_{\mathrm{t}}} \big( M_{3:2048}(t) \big)
\end{equation}

@@

Furthermore:
\begin{equation}
\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_H)_H}_{\mathrm{r}}
\end{equation}

@@

The values stemming from the computation of transactions, specifically the \hyperlink{Transaction Receipt}{transaction receipts, $B_{\mathrm{r}}$}, and that defined through the transaction's \hyperlink{Pi}{state-accumulation function, $\Pi$}, are formalised later in section \ref{sec:statenoncevalidation}.

@@

We define $P(B_H)$ to be the parent block of $B$, formally:
\begin{equation}
P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_H)) = \hyperlink{H p}{H_{\mathrm{p}}}
\end{equation}

The block number is the parent's block number incremented by one:
\begin{equation}
\hyperlink{H i}{H_{\mathrm{i}}} \equiv {{P(H)_H}_{\mathrm{i}}} + 1
\end{equation}

@@

The nonce of a block, $\hyperlink{H n}{H_{\mathrm{n}}}$, must satisfy the relations:
\begin{equation}
n \leqslant \frac{2^{256}}{\hyperlink{H d}{H_{\mathrm{d}}}} \quad \wedge \quad m = \hyperlink{H m}{H_{\mathrm{m}}}
\end{equation}
with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, \hyperlink{H n}{H_{\mathrm{n}}}, \mathbf{d})$.
